---
title: "R Notebook"
output: html_notebook
---

# This is a title

## This is a subtitle

These are my notes:

- Note 1
- Note 2

```{r}

print("hello world")

```

# IF, FOR and WHILE 

```{r}
# if statement

age <- 25

if (age<10){
  print("Get off social media")
} else if (age < 18){
  print("Use with caution")
} else{
  print("Welcome!")
}
```

```{r}
#WHILE LOOP 

x <- 1

while (x<10){
  print(x)
  x <- x + 1
}
```

```{r}
#FOR loop
names <- c("Alastair", "Martin", "Lewis")

for (elephant in names){
  print(elephant)
}
```

# Working with CSVs

```{r}
data <- read.csv("../DATA/gapminder.csv")
data
```
Turning categorical variables into factors...
```{r}
data$country <- as.factor(data$country)
data$region <- as.factor(data$region)
```

```{r}
max(data$year)
min(data$life_expectancy)
mean(data$population)
unique(data$region)
```

```{r}
# REMEMBER [ROWS,COLS]

data[data$life_expectancy > 30,c("population", "life_expectancy")]
```

```{r}
q1 <- data[data$population < 1000000,]
q1
```

```{r}
#List the countries for which gdp_per_capita in 2011 was over 60000, sorted by gdp_per_capita.

q2 <- data[data$gdp_per_capita > 60000 & data$year == 2011,]
q2 <- q2[order(q2$gdp_per_capita,decreasing=TRUE)]
q2
```


```{r}
#Q3 
europe_82 <- data[data$region == "Europe" & data$life_expectancy > 82,]

unique(europe_82$country)
```

1. Create a variable asia_82 that contains the following fields: year, country, gdp_per_capita, life_expectancy for all the countries in Asia region where life expectancy is over 82 that refer to years 2014 and later.
2. List all the countries that can be found in this subset (there should be 3).

```{r}
asia_82 <- data[data$life_expectancy > 82 & data$year >= 2014 & data$region == "Asia", c("year", "country", "gdp_per_capita", "life_expectancy")]

#unique(asia_82$country)
```

# Basic Visualisations

```{r}
# Histogram

hist(data$life_expectancy)

# Adding parameters

hist(data$life_expectancy,
     breaks = 35,
     xlab="Life Exp",
     ylab="Freq",
     col="#CC55FF")

```


```{r}
# Bar Chart

## Aggregating the data
country_regions <- unique(data[,c("region", "country")])
country_regions

counts <- table(country_regions$region)
counts

barplot(counts)
```

```{r}
# Subplots

d1850 <- data[data$year == 1850,]
d1950 <- data[data$year == 1950,]
d2015 <- data[data$year == 2015,]

par(mfrow=c(1,3)) # Creates a 3 by 1 grid for the plot
hist(d1850$life_expectancy)
hist(d1950$life_expectancy)
hist(d2015$life_expectancy)
```


# GGPLOT2

```{r}
# Importing a library
library(ggplot2)
```

```{r}
carData <- ggplot2::mpg
carData

#?mpg
```


```{r}
# Scatter Plot 

ggplot(data=carData) +
  geom_point(mapping=aes(x=displ, y=hwy))

# Adding colours

ggplot(data=carData) +
  geom_point(mapping=aes(x=displ, y=hwy), color="blue", shape=17)

# # Adding colours

ggplot(data=carData) +
  geom_point(mapping=aes(x=displ, y=hwy, color=class), shape=17)

# Changing the size of the point based on a col value

ggplot(data=carData) +
  geom_point(mapping=aes(x=displ, y=hwy, color=class, size=cyl), shape=17)
```

```{r}
# Working with categorical variables

ggplot(data=carData) +
  geom_point(mapping=aes(x=displ, y=hwy)) +
  facet_wrap(~class, nrow=2)

```

```{r}
# Working with 2 categorical variables

ggplot(data=carData)+
  geom_point(mapping=aes(x=displ, y=hwy)) +
  facet_grid(drv~cyl)
```

## Line Graphs

```{r}
# Basic Line Graph
ggplot(data=carData) +
  geom_smooth(mapping=aes(x=displ,y=hwy))

# Splitting the lines
ggplot(data=carData) +
  geom_smooth(mapping=aes(x=displ,y=hwy, linetype=drv))
```


## Combining Graphs

```{r}
# Combining a scatter and a line graph

ggplot(data=carData) +
  geom_point(mapping=aes(x=displ,y=hwy)) +
  geom_smooth(mapping=aes(x=displ,y=hwy))

# Global Mappings
ggplot(data=carData, mapping=aes(x=displ, y=hwy)) +
  geom_point(mapping=aes(color=class))+
  geom_smooth()
```


## Bar Graphs

```{r}
diamonds <- ggplot2::diamonds
diamonds
```


```{r}
# Bar chart 

ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut))

# Filling each bar 

ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut, fill=clarity))

# Splitting the above
ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut, fill=clarity), position="dodge")

# Another view...
ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut, fill=clarity), position="fill")


```
























